---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## 1. Loading and preprocessing the data

I will use `unzip()` and `read.csv()` to access the file, and `dplyr::as_tibble()` to format it.

```{r, message = FALSE}
library(dplyr)
unzip("activity.zip")
temp <- read.csv("activity.csv")
activity <- as_tibble(temp)
```

---

## 2. What is mean total number of steps taken per day?

### 2.1. Calculate total steps per day.

I will use "dplyr" for this first task.

```{r}
library(ggplot2)
daily_steps <- activity %>%
        group_by(date) %>%
        summarise(total_steps = sum(steps))
daily_steps
```

### 2.2. Make a histogram.

I will use "ggplot2" to plot a histogram (remembering that histogram is a frequency plot with a continuous variable - in this case, total steps - organised into bins).

```{r}
ggplot(daily_steps) +
        geom_histogram(mapping = aes(x = total_steps), binwidth = 500)
```

### 2.3. Calculate and report the mean and median steps per day.

With the `daily_steps` tibble I can calculate mean and median values:

```{r}
summarise(daily_steps, mean_steps = mean(total_steps, na.rm = TRUE), median_steps = median(total_steps, na.rm = TRUE))
```

---

## 3. What is the average daily activity pattern?

### 3.1. Time series plot

Find the mean number of steps per time interval using "dplyr", then graph it with "ggplot2".

```{r}
by_interval <- activity %>%
        group_by(interval) %>%
        summarise(mean_steps = mean(steps, na.rm = TRUE))
ggplot(by_interval) +
        geom_line(mapping = aes(x = interval, y = mean_steps))
```

### 3.2. Which 5 minute interval contains the maximum number of steps, on average?

I can subset out the portion of the `by_interval` table containing this.

```{r}
by_interval[which(by_interval$mean_steps == max(by_interval$mean_steps)),]
```
The answer is therefore 835. 

---

## 4. Imputing missing values

### 4.1. Calculate and report total number of missing values.

I can use the `complete.cases()` function: the inverse will give us the total number of rows containing an `NA`.

```{r}
sum(!complete.cases(activity))
```
### 4.2. Devising a strategy for all the missing step values.

I have chosen to use mean values for each interval, given these have been previously calculated and exist in the `by_interval' tibble. 

```{r}
by_interval
```

### 4.3. Creating a new dataset with the missing data filled in.

The following code:

1. Creates a copy of the original data frame, called `activity2`.
2. Converts the data in the "steps" column to numeric.
3. Uses a `for` loop to run through each row of the `activity2` data frame.
4. Uses an `if` argument to identify all the NAs.
    * If an NA is found, this value is then replaced with the mean value for the corresponding interval, as obtained from the `by_interval` tibble.

```{r, cache = TRUE}
activity2 <- as_tibble(temp)
activity2$steps <- as.numeric(activity2$steps)
for(i in 1:nrow(activity2)) {
        if(is.na(activity2[i, "steps"])) {
                activity2[i, "steps"] <- as.numeric(by_interval[which(by_interval$interval == as.integer(activity2[i, "interval"])),2])
        } 
}
activity2
```

### 4.4. Making a histogram of total steps, and reporting mean and median steps for each data set.

Firstly, create a table showing the total steps for the `activity2` data set.

```{r}
daily_steps2 <- activity2 %>%
        group_by(date) %>%
        summarise(total_steps2 = sum(steps))
daily_steps2
```

Merge the `daily_steps` and `daily_steps2` tables together, and keep the merged data frame as a tibble.

```{r}
both_daily_steps <- as_tibble(merge(daily_steps, daily_steps2))
```

Then plot two histograms comparing 


## Are there differences in activity patterns between weekdays and weekends?
